Spatial Model

### 2019-12-13 rmds/poisson_regression.Rmd
work on poisson & spatial, negbinom.stan not in use

where $\lambda_i$ is the relative "risk" of ridership given exposure to covariates $x\beta$ and \theta models extra-Poisson variance.

@@

```{r}
mod_dat <- mod_dat[!is.na(daily_boards) & !is.na(pop_density) & !is.na(emp_density) & !is.na(perc_no_veh) & !is.na(walkability) & !is.na(daily_stops)]

@@

#poisson_fitsimp <- stan(poisson, data = standat1, warmup = 1000, iter = 2000)
poisson_fitsimp <- readRDS("/Users/raven/Documents/honors/honors-project/fits/poisson_fitsimp.RDS")

@@

## Predictions
```{r}
posterior <- extract(poisson_fitsimp)
yrep <- posterior$y_rep
yrep <- yrep[sample(nrow(yrep), 50), ]

pp_check(y, yrep, ppc_dens_overlay) + 
  theme_minimal() +
  xlim(0, 200) +
  labs(title = "Actual versus fitted")
```

@@

Let's see if these residuals are spatially autocorrelated.

@@

@@

One way we could improve this model is by explicitly modeling the extra-Poisson variance (or overdispersion) which may be contributing the the poor fit above. We can add a $\theta$ parameter to address this variation as follows

\begin{align}
Y_i & \sim \text{Poisson}(E_i\lambda_i) \\
\eta_i = log(\lambda_i) &= \mu + \sum_{n=1}^{4}x_n\beta + \theta \\
\theta & \sim N(0, 1)
\end{align}

@@

poisson_theta_fit <- stan(poisson_theta, data = standat1, warmup = 7000, iter = 8000)
saveRDS(poisson_theta_fit, "/Users/raven/Documents/honors/honors-project/stan/poisson_theta_test.stan")

beepr::beep("fanfare")

print(poisson_theta_fit, pars = c("mu", "betas[1]", "betas[2]", "betas[3]", "betas[4]", "theta[5]", "eta[5]", "lambda[5]"), probs=c(0.025, 0.5, 0.975))
launch_shinystan(poisson_theta_fit)
```

```{r}
fit <- rstan::extract(poisson_theta_fit)
test <- rpois(fit$lambda, n = 1513)

ggplot(mod_dat, aes(x=test)) +
  geom_density(color = "red") +
  geom_density(aes(x=daily_boards*365), color = "blue") + xlim(0, 500)
```

So this isn't great but I think it might just need more iterations? This is on the "run overnight" list.

# Testing total annual rides

```{r}
standat1 <- list(y = y*365, E = E, x = x, K = K, N = N)

poisson_fit_year <- stan(poisson, data = standat1, warmup = 1000, iter = 2000)
#poisson_fitsimp <- readRDS("/Users/raven/Documents/honors/honors-project/fits/poisson_fitsimp.RDS")
beepr::beep("fanfare")
print(poisson_fit_year, pars = c("mu", "betas[1]", "betas[2]", "betas[3]", "betas[4]"), probs=c(0.025, 0.5, 0.975))
```

```{r}
fityear <- rstan::extract(poisson_fit_year)
testyear <- rpois(fityear$lambda, n = 1513)

ggplot(mod_dat, aes(x=testyear)) +
  geom_density(color = "red") +
  geom_density(aes(x=daily_boards*365), color = "blue") + xlim(0, 3000)
```

# Testing negative

```{r}
negbinom <- "/Users/raven/Documents/honors/honors-project/stan/negbinom.stan"

nb_fit <- stan(negbinom, data = standat1, warmup = 1000, iter = 2000, control = list(adapt_delta = 0.99))

print(nb_fit, pars = c("mu", "betas[1]", "betas[2]", "betas[3]", "betas[4]", "eta[5]", "lambda[5]", "scale"), probs=c(0.025, 0.5, 0.975))

launch_shinystan(nb_fit)
```

>>>>

where $\lambda_i$ is the relative "risk" of ridership given exposure to covariates $x\beta$.

@@

```{r}
mod_dat <- mod_dat[!is.na(daily_boards_per_stop) & !is.na(pop_density) & !is.na(emp_density) & !is.na(perc_no_veh) & !is.na(walkability) & !is.na(daily_stops)]

@@

#poisson_fitsimp <- stan(poisson, data = standat1, warmup = 1000, iter = 2000)
#saveRDS(poisson_fitsimp, "/Users/raven/Documents/honors/honors-project/fits/poisson_fitsimp.RDS")

poisson_fitsimp <- readRDS("/Users/raven/Documents/honors/honors-project/fits/poisson_fitsimp.RDS")

@@

## Predictions
```{r}
posterior <- rstan::extract(poisson_fitsimp)
yrep <- matrix(ncol = 1513, nrow = 100)

for(i in 1:100){
  yrep[i, ] <- rpois(1513, posterior$lambda*E)
}

pp_check(y, yrep, ppc_dens_overlay) + 
  theme_minimal() +
  xlim(0, 200) +
  labs(title = "Actual versus fitted") + xlim(0, 100)
```

@@

Let's see if these residuals are spatially autocorrelated:

@@

According to the Moran simulation, they're not spatially clustered, but we still haven't found a good way to measure "linear" clustering.

@@

One way we could improve this model is by explicitly modeling the extra-Poisson variance (or overdispersion) which may be contributing the the poor fit above. We can add a $\theta$ parameter (scaled by $\sigma$, following the bym2 model) to address this variation as follows

\begin{align}
Y_i & \sim \text{Poisson}(E_i\lambda_i) \\
\eta_i = log(\lambda_i) &= \mu + \sum_{n=1}^{4}x_n\beta + \theta*\sigma \\
\theta & \sim N(0, 1)
\end{align}

@@

#poisson_theta_fit <- stan(poisson_theta, data = standat1, warmup = 1000, iter = 2000)
#saveRDS(poisson_theta_fit, "/Users/raven/Documents/honors/honors-project/fits/poisson_theta_test.RDS")

poisson_theta_fit <- readRDS("/Users/raven/Documents/honors/honors-project/fits/poisson_theta_test.RDS")

print(poisson_theta_fit, pars = c("mu", "betas[1]", "betas[2]", "betas[3]", "betas[4]", "theta[5]", "eta[5]", "lambda[5]"), probs=c(0.025, 0.5, 0.975))

launch_shinystan(poisson_theta_fit)
```

This also has a lot of autocorrelation in the chains, needs to be run for more iterations.

```{r}
posterior2 <- rstan::extract(poisson_theta_fit)
yrep2 <- matrix(ncol = 1513, nrow = 100)

for(i in 1:100){
  yrep2[i, ] <- rpois(1513, posterior2$lambda[i,]*E)
}

pp_check(y, yrep2, ppc_dens_overlay) + 
  theme_minimal() +
  xlim(0, 200) +
  labs(title = "Actual versus fitted") + xlim(0, 100)
```

The next improvement would be to determine roughly how much of the variation currently modeled by $\theta * \sigma$ is due to unmeasured covariates vs pure overdispersion vs spatial variance.

### 2019-12-13 rmds/spatial_mod.Rmd
work on poisson & spatial, negbinom.stan not in use

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(INLA)
library(rstan)
library(shinystan)
library(data.table)
library(ggplot2)
library(rgdal)
library(sf)

@@

bym2 <- "/Users/raven/Documents/honors/honors-project/stan/bym2.stan"

#bym2_fit1 <- stan(bym2, data = standat1, warmup = 5000, iter = 6000, verbose = T)
#saveRDS(bym2_fit1, "/Users/raven/Documents/honors/honors-project/fits/bym2_fit1.RDS")

bym2_fit1 <- readRDS("/Users/raven/Documents/honors/honors-project/fits/bym2_fit1.RDS")
```

```{r}
print(bym2_fit1, pars=c("beta0", "betas[1]", "rho", "sigma", "logit_rho", "mu[5]", "phi[5]", "theta[5]"), probs=c(0.025, 0.5, 0.975))
#launch_shinystan(bym2_fit1)
```

I'm not sure if the problem with this is that I didn't run it enough, it's parameterized poorly, etc. Actually only took ~10 minutes to run.

# All four covariates, Ramsey

Using not enough iterations right now so I can see if it works
```{r}
x2 <- matrix(ncol = 4, c(spat_dat$pop_density, spat_dat$emp_density, spat_dat$perc_no_veh, spat_dat$walkability))
K2 <- 4

@@

```{r}
#launch_shinystan(bym2_fit2)
print(bym2_fit2, pars=c("beta0", "betas[1]", "betas[2]", "betas[3]", "betas[4]", "rho", "sigma", "logit_rho", "mu[5]", "phi[5]", "theta[5]"), probs=c(0.025, 0.5, 0.975))
```

Everything except $\rho$ seems to be alright..

>>>>

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
library(INLA)
library(rstan)
library(shinystan)
library(data.table)
library(ggplot2)
library(rgdal)
library(sf)
library(tigris)
library(spdep)
library(bayesplot)

@@

# bym2 <- "/Users/raven/Documents/honors/honors-project/stan/bym2.stan"
# 
# #bym2_fit1 <- stan(bym2, data = standat1, warmup = 5000, iter = 6000, verbose = T)
# #saveRDS(bym2_fit1, "/Users/raven/Documents/honors/honors-project/fits/bym2_fit1.RDS")
# 
# bym2_fit1 <- readRDS("/Users/raven/Documents/honors/honors-project/fits/bym2_fit1.RDS")
```

<!-- ```{r} -->
<!-- print(bym2_fit1, pars=c("beta0", "betas[1]", "rho", "sigma", "logit_rho", "mu[5]", "phi[5]", "theta[5]"), probs=c(0.025, 0.5, 0.975)) -->
<!-- #launch_shinystan(bym2_fit1) -->
<!-- ``` -->

<!-- I'm not sure if the problem with this is that I didn't run it enough, it's parameterized poorly, etc. Actually only took ~10 minutes to run.  -->

# All four covariates, Ramsey County

```{r}
x2 <- matrix(ncol = 4, c(spat_dat$pop_density, spat_dat$emp_density, spat_dat$perc_no_veh, spat_dat$walkability))
K2 <- 4

@@

```{r}
launch_shinystan(bym2_fit2)
print(bym2_fit2, pars=c("beta0", "betas[1]", "betas[2]", "betas[3]", "betas[4]", "rho", "sigma", "logit_rho", "mu[5]", "phi[5]", "theta[5]"), probs=c(0.025, 0.5, 0.975))
```

Some of these terms haven't fully converged but I've only run the chains for 2000 iterations. Using `shinystan`, I can see that there's a lot of autocorrelation in the chains. I think if there are no other red flags I can run this for many more iterations for a good result.

```{r}
posteriorbym2 <- rstan::extract(bym2_fit2)
yrepbym2 <- matrix(ncol = 357, nrow = 100)

for(i in 1:100){
  yrepbym2[i, ] <- rpois(357, posteriorbym2$mu[i, ]) # reparameterize
}

pp_check(y, yrepbym2, ppc_dens_overlay) + 
  theme_minimal() +
  xlim(0, 200) +
  labs(title = "Actual versus fitted") + xlim(0, 100)
```

This fit should also be improved once I finish variable selection.
